Improve Taxonomy Assignment

### 2022-02-24 reports/01_taxonomy_improved.Rmd
initial commit

---
title: "Improve Taxonomy Assignment"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)

#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
```

@@

#############
# LODA DATA #
#############
paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region

n <- c("boston_r1", "boston_r2", "Sthlm")
ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

@@

######################
# DOWNLOAD DATABASES #
######################
db_dir <- "../../data/database/"
if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }

@@

walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
system("gunzip ../../data/database/*.gz")

@@

@@

write.csv(Seq_ID_key, file=paste0("../../results/Seq_ID_key.csv"), quote=FALSE, row.names=FALSE)

@@

# Create Query files
write.fasta(Seq_ID_key, paste0("../../results/BVAB_Query.fasta"))
write.fasta(unnasigned_0, paste0("../../results/optivag_Query.fasta"))
```

@@

```{r BLASTn search}
# Make uniqe identifyers in the BVAB file
fa = readDNAStringSet(paste0(db_dir,"BVAB_rRNA_database.fa"))
ID <-  make.unique(paste0(str_match(names(fa), "^BVAB._"),"0"), sep = "")
names(fa) = paste(ID, names(fa))
writeXStringSet(fa, "../../data/database/BVAB_db.fasta")

for(name in c("BVAB", "optivag")) { 
  db_dir     = "../../data/database/"
  result_dir = "../../results/"
  if (isFALSE(is_empty(list.files(path = db_dir, pattern = paste0(name,".+\\.nhr"))))) {
    message("using database provided in path:")
    message(paste0(db_dir, name, "_reference.db"))
  } else {
    system2(command = "makeblastdb", 
                          args = c("-in", paste0(db_dir, name, "_db.fasta"), 
                                  "-out", paste0(db_dir, name, "_reference.db"), 
                                  "-parse_seqids", 
                                  "-dbtype", "nucl"),
                         wait = TRUE,
                         stdout = TRUE)
  }
  blast_out <- system2(command = "blastn", 
                       args = c("-db", paste0(db_dir, name, "_reference.db"), 
                                "-query", paste0(result_dir, name, "_Query.fasta"), 
                                #"-out", paste0(result_dir, name, "_hits.tab"),
                                "-outfmt", 6, 
                                "-perc_identity", 99.5,
                                "-qcov_hsp_perc", 90),
                       wait = TRUE,
                       stdout = TRUE) %>%
  as_tibble() %>% 
  separate(col = value, sep = "\t", convert = TRUE,
           into = c("qseqid","sseqid","pident","length","mismatch","gapopen",
                    "qstart","qend","sstart","send","evalue","bitscore") )

  write_tsv(blast_out, paste0("../../results/", name, "_hits.tsv"))
}

@@

```{r Load BLAST hits, message=FALSE}
blast_OptiVag <- read_table2(paste0("../../results/OptiVag_hits.tsv"))
blast_BVAB <- read_table2(paste0("../../results/BVAB_hits.tsv"))

@@

write.csv(new_taxtable, file=paste0("../../data/SeqID_to_taxa.csv"), quote=FALSE, row.names=FALSE)
```

@@

write.csv(unnasigned_final, 
          file=paste0("../../results/Unnasigned_tax_report.csv"), quote=FALSE, row.names=FALSE)

@@

#### ASV with all controls and replicate samples removed ####
ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3_B1", "ASV_tissue_V3_B1", 
               "ASV_CVL_V2_B2", "ASV_tissue_V3_B2", "ASV_CVL_V2_S" )
ASV_list <- map(ASV, ~ .x %>%
                  filter(grepl("yes", .$Include)) %>%
                  group_split(Sample_type, Visit, .keep = T)
            ) %>%
            flatten() %>%
            set_names(ASV_names) %>%
            map(., ~ .x %>%
                select(Sequence, ID, Reads) %>%
                pivot_wider(., ,names_from=ID, values_from=Reads) %>%
                left_join(., new_taxtable, by="Sequence") 
                )

@@

```{r Save ASV files}
imap(ASV_list, 
     ~write.csv(.x, file=paste0("../../data/",.y,".csv"), quote=FALSE, row.names=F) )

saveRDS(ASV, paste0(PATH,"results/","ASV_wCtrl_long_format.RDS"))
```

>>>>

---
title: "Improve Taxonomy Assignment"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)

# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
```

@@

#########
# PATHS #
#########
result_dir <- "../results/01_taxonomy_improved_output/"
db_dir <- "../resources/OptiVag_BVAB_database/"

#############
# LODA DATA #
#############
#### new ####
paths_p <- c('../results/00_dada2_output/ps_objects/phyloseq_Boston_run1.RDS', # 108 Luminal samples, 1 Tissue sample
             '../results/00_dada2_output/ps_objects/phyloseq_Boston_run2.RDS'  # 92 Tissue samples
             )
##### old ####
paths_p <- c('phyloseq_boston_r1.RDS',   # 108 Luminal samples, 1 Tissue sample
             'phyloseq_boston_r2.RDS')   # 92 Tissue samples

n <- c("boston_r1", "boston_r2")
ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

#########################
# SELECT SAMPLES TO USE #
#########################
mapping <- c("../data/Mapping/Mapping_info_Boston_run1.csv", "../data/Mapping/Mapping_info_Boston_run2.csv")
m <- mapping %>% 
  map(., ~read_csv(file = .x, col_names = TRUE)) %>% 
  map(., ~pull(.x, "Sample_ID"))  %>% 
  flatten()
ps_list <- map(ps_list, ~subset_samples(.x, Sample_ID %in% m))
#map2(ps_list, paths_p, ~saveRDS(.x, file = paste0("../data/", .y)))

@@

######################
# DOWNLOAD DATABASES #
######################
if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }

@@

walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
system("gunzip ../resources/OptiVag_BVAB_database/*.gz")

@@

```{r test identity, eval=FALSE, include=FALSE}
paths_p <- c('../results/00_dada2_output/ps_objects/phyloseq_Boston_run1.RDS', # 108 Luminal samples, 1 Tissue sample
             '../results/00_dada2_output/ps_objects/phyloseq_Boston_run2.RDS'  # 92 Tissue samples
             )
paths_p <- c('../../Gabriella_repo/data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
             '../../Gabriella_repo/data/phyloseq_boston_r2.RDS')       # CVL V2, only v4 region

n <- c("boston_r1", "boston_r2")
ps_list_new <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

seqtabs_ <- map(ps_list_new, ~data.frame(.x@otu_table@.Data))
taxtabs_ <- map(ps_list_new, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
sample_info_ <- map(ps_list_new, ~as_tibble(sample_data(.x), rownames = 'SampleID'))

int <- intersect(colnames(seqtabs[[1]]), colnames(seqtabs_[[1]]))
setdiff(colnames(seqtabs[[1]]), colnames(seqtabs_[[1]]))
setdiff(colnames(seqtabs_[[1]]), colnames(seqtabs[[1]]))

old <- seqtabs_[[1]][row.names(seqtabs[[1]]), int]
new <- seqtabs[[1]][, int]

all_equal(old, new)

old[1,1:5]
new[1,1:5]

head(seqtabs)

@@

if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }
write.csv(Seq_ID_key, file=paste0(result_dir,"Seq_ID_key.csv"), quote=FALSE, row.names=FALSE)

@@

# Create Query files
write.fasta(Seq_ID_key, paste0(result_dir,"BVAB_Query.fasta"))
write.fasta(unnasigned_0, paste0(result_dir,"optivag_Query.fasta"))
```

@@

```{r BLASTn search}
# Make uniqe identifyers in the BVAB file
fa = readDNAStringSet(paste0(db_dir,"BVAB_rRNA_database.fa"))
ID <-  make.unique(paste0(str_match(names(fa), "^BVAB._"),"0"), sep = "")
names(fa) = paste(ID, names(fa))
writeXStringSet(fa, paste0(db_dir,"BVAB_db.fasta"))

for(name in c("BVAB", "optivag")) {
  if (isFALSE(is_empty(list.files(path = db_dir, pattern = paste0(name,".+\\.nhr"))))) {
    message("using database provided in path:")
    message(paste0(db_dir, name, "_reference.db"))
  } else {
    system2(command = "makeblastdb", 
                          args = c("-in", paste0(db_dir, name, "_db.fasta"), 
                                  "-out", paste0(db_dir, name, "_reference.db"), 
                                  "-parse_seqids", 
                                  "-dbtype", "nucl"),
                         wait = TRUE,
                         stdout = TRUE)
  }
  blast_out <- system2(command = "blastn", 
                       args = c("-db", paste0(db_dir, name, "_reference.db"), 
                                "-query", paste0(result_dir, name, "_Query.fasta"), 
                                #"-out", paste0(result_dir, name, "_hits.tab"),
                                "-outfmt", 6, 
                                "-perc_identity", 99.5,
                                "-qcov_hsp_perc", 90),
                       wait = TRUE,
                       stdout = TRUE) %>%
  as_tibble() %>% 
  separate(col = value, sep = "\t", convert = TRUE,
           into = c("qseqid","sseqid","pident","length","mismatch","gapopen",
                    "qstart","qend","sstart","send","evalue","bitscore") )

  write_tsv(blast_out, paste0(result_dir, name, "_hits.tsv"))
}

@@

```{r Load BLAST hits, message=FALSE}
blast_OptiVag <- read_table2(paste0(result_dir,"optivag_hits.tsv"))
blast_BVAB <- read_table2(paste0(result_dir, "BVAB_hits.tsv"))

@@

write.csv(new_taxtable, file=paste0(result_dir,"SeqID_to_taxa.csv"), quote=FALSE, row.names=FALSE)
```

@@

write.csv(unnasigned_final, 
          file=paste0(result_dir,"Unnasigned_tax_report.csv"), quote=FALSE, row.names=FALSE)

@@

#### ASV with all controls and replicate samples removed ####
# ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3_B1", "ASV_tissue_V3_B1", 
#                "ASV_CVL_V2_B2", "ASV_tissue_V3_B2" )
ASV_names <- c("ASV_CVL_V3_B1", "ASV_tissue_V3_B1", "ASV_tissue_V3_B2" )
ASV_list <- map(ASV, ~ .x %>%
                  filter(grepl("yes", .$Include)) %>%
                  group_split(Sample_type, Visit, .keep = T)
            ) %>%
            flatten() %>%
            set_names(ASV_names) %>%
            map(., ~ .x %>%
                select(Sequence, ID, Reads) %>%
                pivot_wider(., ,names_from=ID, values_from=Reads) %>%
                left_join(., new_taxtable, by="Sequence") 
                )

@@

```{r Save ASV files}
imap(ASV_list, 
     ~write.csv(.x, file=paste0(result_dir,.y,".csv"), quote=FALSE, row.names=F) )

saveRDS(ASV, paste0(result_dir,"ASV_wCtrl_long_format.RDS"))
```
